Appendix to 'Detecting social information in a dense database of infants’ natural

### 2021-05-25 paper/preprint/paper_revised_submission_april2021_apa.Rmd

```{r}
# Read in data 
## Face/hand detections by video
load(here::here('data/preprocessed_data_2021/all_vid_data_from_bbs_all_detections.RData')) # all detections
```

@@

```{r}
# gold sample annotatios and detedctinos for these specific frames
load(here::here('data/preprocessed_data_2021/gold_sample_annotations2020-01-31.RData')) # human annotations
load(here::here('data/preprocessed_data_2021/gold_sample_from_bbs_2021_all_detections.RData'))

# all_gold_sample_frames_op_hc = load(here::here('data/preprocessed_data_2021/gold_sample_from_bbs_2021_high_conf_detections.RData')) 
```

@@

```{r eval=FALSE,include=FALSE}
## for high confidence detections
# gold_sample_hc <- gold_sample %>%
#   select(vid_name, frame, face_present_ketan, hand_present_ketan) %>%
#   mutate(face_present_ketan = as.logical(face_present_ketan), hand_present_ketan = as.logical(hand_present_ketan)) %>%
#   right_join(all_gold_sample_frames_op_hc) %>% # HIGH CONF DF
#   mutate(face_eval_ketan = evaluate_detector(face_present_ketan, face_detected), hand_eval_ketan = evaluate_detector(hand_present_ketan, hand_detected))
# 
#  
# face_performance_hc = return_prf_short(gold_sample_hc$face_eval_ketan)
# hand_performance_hc = return_prf_short(gold_sample_hc$hand_eval_ketan)
```

@@

```{r} 
### Load turk hand annotations with all bounding boxes (multiple dets per frame)
load(here::here('data/preprocessed_data_2021/hand_annotations_2020-01-29.RData'))

child_hands <- hand_annotations %>%
  filter(label=="Child hand")

adult_hands <- hand_annotations %>%
  filter(label=="Adult hand")
```

```{r}
### plot centers of the bounding boxes made for child and adult hands
child_hand_plot <- ggplot(child_hands, aes(x=center_x, y=center_y)) +
  geom_point(alpha=.1) + 
  stat_density_2d(aes(fill = ..level..), geom="polygon", alpha=.8) +
  coord_fixed(ratio=1) +
  ggtitle('A. Child hand density') +
  ylim(0,480) +
  xlim(0,640) +
  ylab('') +
  xlab('') +
  theme_few(base_size=6) +
  theme(legend.position="none", axis.text.x=element_blank(), axis.ticks.x=element_blank(), axis.text.y=element_blank(), axis.ticks.y=element_blank()) +
  scale_y_reverse()
##
adult_hand_plot <- ggplot(adult_hands, aes(x=center_x, y=center_y)) +
  geom_point(alpha=.1) + 
  stat_density_2d(aes(fill = ..level..), geom="polygon", alpha=.8) +
  coord_fixed(ratio=1) +
  ylim(0,480) +
  xlim(0,640) +
  scale_y_reverse() +
  ylab('') +
  xlab('') +
  theme_few(base_size=6) +
    theme(legend.position="none", axis.text.x=element_blank(), axis.ticks.x=element_blank(), axis.text.y=element_blank(), axis.ticks.y=element_blank()) +
  ggtitle('B. Adult hand density')
```

@@

```{r handLocation, fig.env="figure", fig.pos = "h", fig.align = "center", fig.width=5, fig.height=3, fig.cap = "Density estimates for the child (left) and adult (right) hands that were detected in the 24K frame random gold set; each dot represents the center of a bounding box made by an adult participant. Brighter values indicate more detections."}
cowplot::plot_grid(child_hand_plot, adult_hand_plot, nrow=1)
```

We suspected that this was in part because children’s own hands were often in view of the camera and unconnected to a pose—-a notoriously challenging detection problem [@bambach2015lending]. To assess this possibility, we obtained human annotations for the entire subsample of 9051 frames in which a hand was detected; participants (recruited via AMT) were asked to draw bounding boxes around children’s and adult’s hands. Overall, we found that `r round(mean(hands_missed$child_hand_seg),2)*100`% of missed hand detections were of child hands. When frames with children's hands were removed from the gold set, recall did improve somewhat to `r round(hand_performance_adults[3],2)`. We also observed that children’s hands tended to appear in the lower half of the frames; heatmaps of the bounding boxes obtained from these annotations can be seen in Figure \ref{fig:handLocation}.

@@

```{r}
#### Make longer format for plotting
all_detections <- face_hand_by_age %>%
  filter(num_frames_total > 2000) %>% # eliminate small data point that skews scaling
  select(prop_faces, prop_hands, num_frames_total, age_day_bin, child_id) %>%
  pivot_longer(cols = c(prop_faces, prop_hands), names_to = "region", values_to = "prop") %>%
  mutate(approach = "uncropped",
         region = ifelse(region == "prop_faces","Faces","Hands"))

centered <- face_hand_by_age %>%
  filter(num_frames_total > 2000) %>% # eliminate small data point that skews scaling
  select(prop_faces_center, prop_hands_center, num_frames_total, age_day_bin, child_id) %>%
  gather(region, prop, prop_faces_center, prop_hands_center) %>%
  mutate(approach = "centered",
         region = ifelse(region == "prop_faces_center","Faces","Hands"))
```

```{r FacesAndHands, fig.env="figure*", fig.pos = "h", fig.align = "center", fig.width=7, fig.height=3, fig.cap = "Proportion of faces and hands seen as a function of age for each child (A, S, and Y). Data are binned by each week that the videos were filmed and scaled by the number of frames in that age range. Dashed lines show estimated trend lines from proportion of faces/hands in view when detections are restricted to the center FOV, reducing the contribution of children's own hands."}
ggplot(all_detections, 
       aes(x=age_day_bin, y=prop, 
           size=log10(num_frames_total),
           col=region)) +
  geom_point(alpha=.2) +
  geom_smooth(span=10, aes(weight = num_frames_total), show.legend = FALSE) + 
  geom_smooth(data = centered, span=10, aes(weight = num_frames_total), show.legend = TRUE,
              lty = 2, span=10, se = FALSE) +
  ylab('Proportion Detections') + 
  xlab('Age (Months)') +
  ylim(0,.5) +
  facet_grid(.~child_id) + 
  theme_few(base_size=10) +
  ggthemes::scale_color_solarized(name = "") + 
  theme(legend.text=element_text(size=10)) +
  guides(size = FALSE) +
  theme(legend.position="bottom") 
```

We analyzed the social information in view across the entire dataset, looking specifically at the proportions of faces and hands detected for each child.^[All analyses and preprocessed data files for this paper are available at https://tinyurl.com/detecting-social-info] Data from videos were binned according to the age of the child (in weeks). First, we saw that the proportion of faces in view showed a moderate decrease across this age range (see Figure \ref{fig:FacesAndHands}), in keeping with prior findings [@Fausey2016]; in contrast, we did not observe an increase in the proportion of hands in view. These effects were quantified with two separate linear mixed-models (see Tables 1 & 2).^[Face/hand detections were binned across each week of filming. Participant's age was converted into months and centered for these analyses. Random slopes for the effect of age by child led to a singular fit and were removed from both analyses; see full model specification in accompanying codebase.]

However, the most striking result from these analyses is a much overall greater proportion of hands in view than have previously been reported [@Fausey2016]. We found this to be true across all ages, in all three children, and regardless of whether we analyzed human annotations (on the 24K random subset, see dotted lines in Appendix Figure \ref{fig:goldSetSanity}) or OpenPose annotations on the entire dataset (see solid lines in Figure \ref{fig:FacesAndHands}). This is notable especially given that OpenPose showed relatively low recall for hands, indicating that this may be an underestimate of the proportion of hands in view. In fact, analysis the human annotations underscores revealed a much higher proportion of hands relative to faces than the automated annotations.

@@

<!-- FIXME - references not rendering for -->
As we found that children's hands tended to occur in the lower visual field (see Figure \ref{fig:handLocation}), we thus re-analyzed the entire dataset while restricting our analysis to the center field of view, decreasing the proportion of hand detections from `r round(sum(all_vid_data$num_hands)/sum(all_vid_data$num_frames),2)*100`% to `r round(sum(all_vid_data$num_hands_center,na.rm=TRUE)/sum(all_vid_data$num_frames),2)*100`%, but only decreased face detections from `r round(sum(all_vid_data$num_faces)/sum(all_vid_data$num_frames),2)*100`% to `r round(sum(all_vid_data$num_faces_center,na.rm=TRUE)/sum(all_vid_data$num_frames),3)*100`%. 
This cropping likely removed both the majority of detections of children's own hands but also some detections of adult hands (see Figure \ref{fig:handLocation}), especially as OpenPose was biased to miss children's hands when they were in view. 
Nonetheless, within this modified field of view, we still observed more hand detections than face detections (see dashed lines in Figure \ref{fig:FacesAndHands}). 
We also still found a higher proportion of hands in view relative to faces when excluding any frames containing child hand's from the human annotated gold sample (see Appendix Figure\ref{fig:goldSetSanity}).

@@

```{r}
face_hand_detailed_plot <- face_hand_by_age %>%
  gather(region, prop, prop_full_faces, prop_faces, prop_hands, prop_faces_and_hands) %>%
  mutate(approach = "uncropped",
         region = case_when(region == "prop_faces" ~"All faces",
                            region == "prop_full_faces" ~ "Full faces (eyes, nose, mouth visible)",
                            region == "prop_hands" ~ "All hands",
                            region == "prop_faces_and_hands" ~ "Both hands and faces visible"),
         child_name = case_when(child_id =='A' ~ "Child A",
                      child_id == "S" ~ "Child S",
                      child_id == "Y" ~ "Child Y")) %>%
  filter(num_frames_total > 5000) # skews scaling of detections to have this 1 sparse age bin

```{r faceHandDetailed, fig.env="figure*", fig.pos = "h", fig.align = "center", fig.width=7, fig.height=3, fig.cap = "Proportion of face, hand, full face (i.e., nose, mouth, and nose visible), and joint face/hand detections as a function of age for each child in the dataset. Data are binned by each week that the videos were filmed and scaled by the number of frames in that age range." }
ggplot(face_hand_detailed_plot, 
       aes(x=age_day_bin, y=prop, 
           col=region)) +
  geom_point(alpha=.1, aes(size=log10(num_frames_total)), show.legend= FALSE) +
  geom_smooth(span=10, aes(weight = num_frames_total), show.legend = TRUE) + 
  ylab('Proportion Detections') + 
  xlab('Age (Months)') +
  ylim(0,.5) +
  facet_grid(.~child_name) + 
  theme_few(base_size=12) +
  scale_color_manual(values = c( "#2048d6","#d62020","#751098","#142c83"), name = "") +
  scale_size_continuous(name = "") +
  guides(size = FALSE) +
  theme(legend.text=element_text(size=8)) +
  theme(legend.position="bottom") 
```

@@

Finally, we also examined whether there were changes in where faces tended to appear in the camera's (and hence, by proxy, the child's) field of view. 
As expected, faces tended to be located towards the upper field of view, while views of hands were more centrally distributed (see Appendix, Figure XX for average density distributions). 
However, we also found evidence that older children tended to see more faces in more variable positions than younger children. 
Specifically, we examined how variable the horizontal and vertical coordinates were of the faces in the infant view. 
To do so, we calculated the coefficient of variation of the horizontal (x) and vertical (y) positions of centers of the faces detected by OpenPose (see Figure \ref{fig:faceDetailed}B), and examined changes across age. 
Faces tended to be more variable in the vertical than their horizontal position (see Figure \ref{fig:faceDetailed}B). 
We also found that as children got older, they tended to see faces that varied more in their horizontal -- but not their vertical position -- suggesting that older children might be more likely to see more smaller faces in their periphery (see Figure \ref{fig:faceDetailed}B).

>>>>

```{r}
# Read in data 
## Face/hand detections by video
load(file=here::here('data/preprocessed_data_2021/all_vid_data_from_bbs_all_detections.RData')) # all detections
```

@@

```{r}
# gold sample annotations for these specific frames
load(here::here('data/preprocessed_data_2021/gold_sample_annotations2020-01-31.RData')) # human annotations
```

```{r}
# load high confidence detections for gold sample frames
load(here::here('data/preprocessed_data_2021/gold_sample_from_bbs_2021_high_conf_detections.RData'))
all_gold_sample_frames_op_hc = all_gold_sample_frames_op
```

```{r}
# load all detections for gold sample frames
load(here::here('data/preprocessed_data_2021/gold_sample_from_bbs_2021_all_detections.RData')) 
```

@@

```{r eval=FALSE,include=FALSE}
## for high confidence detections
gold_sample_hc <- gold_sample %>%
  select(vid_name, frame, face_present_ketan, hand_present_ketan) %>%
  mutate(face_present_ketan = as.logical(face_present_ketan), hand_present_ketan = as.logical(hand_present_ketan)) %>%
  right_join(all_gold_sample_frames_op_hc) %>% # HIGH CONF DF
  mutate(face_eval_ketan = evaluate_detector(face_present_ketan, face_detected), hand_eval_ketan = evaluate_detector(hand_present_ketan, hand_detected))

face_performance_hc = return_prf_short(gold_sample_hc$face_eval_ketan)
hand_performance_hc = return_prf_short(gold_sample_hc$hand_eval_ketan)
```

@@

@@

We suspected that this was in part because children’s own hands were often in view of the camera and unconnected to a pose—-a notoriously challenging detection problem [@bambach2015lending]. To assess this possibility, we obtained human annotations for the entire subsample of 9051 frames in which a hand was detected; participants (recruited via AMT) were asked to draw bounding boxes around children’s and adult’s hands. Overall, we found that `r round(mean(hands_missed$child_hand_seg),2)*100`% of missed hand detections were of child hands. When frames with children's hands were removed from the gold set, recall did improve somewhat to `r round(hand_performance_adults[3],2)`. We also observed that children’s hands tended to appear in the lower half of the frames; heatmaps of the bounding boxes obtained from these annotations can be seen in Appendix Figure \ref{fig:handLocation}.

@@

```{r}
#### Make longer format for plotting
all_detections <- face_hand_by_age %>%
  filter(num_frames_total > 5000) %>% # eliminate small data point that skews scaling
  select(prop_faces, prop_hands, num_frames_total, age_day_bin, child_id) %>%
  pivot_longer(cols = c(prop_faces, prop_hands), names_to = "region", values_to = "prop") %>%
  mutate(approach = "uncropped",
         region = ifelse(region == "prop_faces","All faces","All hands"),
         child_name = case_when(child_id =='A' ~ "Child A",
                      child_id == "S" ~ "Child S",
                      child_id == "Y" ~ "Child Y"))

centered <- face_hand_by_age %>%
  filter(num_frames_total > 5000) %>% # eliminate small data point that skews scaling
  select(prop_faces_center, prop_hands_center, num_frames_total, age_day_bin, child_id) %>%
  gather(region, prop, prop_faces_center, prop_hands_center) %>%
  mutate(approach = "centered",
         region = ifelse(region == "prop_faces_center","Faces in center FOV","Hands in center FOV"),
         child_name = case_when(child_id =='A' ~ "Child A",
                      child_id == "S" ~ "Child S",
                      child_id == "Y" ~ "Child Y"))

detailed <- face_hand_by_age %>%
  filter(num_frames_total > 5000) %>%
  select(prop_full_faces, prop_faces_and_hands, num_frames_total, age_day_bin, child_id) %>%
  gather(region, prop, prop_full_faces, prop_faces_and_hands) %>%
  mutate(approach = "detailed",
         region = case_when(region == "prop_full_faces" ~ "Full faces",
                            region == "prop_faces_and_hands" ~ "Both hands & faces"),
         child_name = case_when(child_id =='A' ~ "Child A",
                      child_id == "S" ~ "Child S",
                      child_id == "Y" ~ "Child Y")) 
```

```{r}
first_panel <- ggplot(all_detections, 
       aes(x=age_day_bin, y=prop, 
           size=log10(num_frames_total),
           col=region)) +
  geom_point(alpha=.1) +
  geom_smooth(span=10, aes(weight = num_frames_total), show.legend = TRUE) + 
  ylab('Proportion Detections') + 
  xlab('Age (Months)') +
  ylim(0,.5) +
  facet_grid(cols = vars(child_name)) + 
  theme_few(base_size=10) +
  ggthemes::scale_color_solarized(name = "") + 
  theme(legend.text=element_text(size=10)) +
  guides(size = FALSE) +
  theme(legend.position="right")

second_panel <- ggplot(centered, 
       aes(x=age_day_bin, y=prop, 
           size=log10(num_frames_total),
           col=region)) +
  geom_point(alpha=.1) +
  geom_smooth(span=10, aes(weight = num_frames_total),linetype = 'dashed', show.legend = TRUE) + 
  ylab('Proportion Detections') + 
  xlab('Age (Months)') +
  ylim(0,.5) +
  facet_grid(cols = vars(child_name)) + 
  theme_few(base_size=10) +
  ggthemes::scale_color_solarized(name = "") + 
  theme(legend.text=element_text(size=10)) +
  guides(size = FALSE) +
  theme(legend.position="right")

third_panel <- ggplot(detailed, 
       aes(x=age_day_bin, y=prop, 
           size=log10(num_frames_total),
           col=region)) +
  geom_point(alpha=.1) +
  geom_smooth(span=10, aes(weight = num_frames_total), show.legend = TRUE) + 
  ylab('Proportion Detections') + 
  xlab('Age (Months)') +
  ylim(0,.5) +
  facet_grid(cols = vars(child_name)) + 
  theme_few(base_size=10) +
  scale_color_manual(values = c("#751098","#142c83"), name = "") +
  theme(legend.text=element_text(size=10)) +
  guides(size = FALSE) +
  theme(legend.position="right")

```{r faceHandDetailed, fig.env="figure*", fig.pos = "h", fig.align = "center", fig.width=7, fig.height=7, fig.cap = "Proportion of frames with (A).All face and hand detections,  B.Face/hand detections that fell within the center field-of-view (reducing the contribution of children's own hands) and (C) Face detections that were full faces (e.g., eyes, nose, and mouth all visible) and that co-occurred with hands, plotted as a function of age for each child (A, S, and Y). Data are binned by each week that the videos were filmed and scaled by the number of frames in that age range."}

cowplot::plot_grid(first_panel, second_panel, third_panel, labels='AUTO', nrow=3, align = "v", axis = "b")
```

We analyzed the social information in view across the entire dataset, looking specifically at the proportions of faces and hands detected for each child.^[All analyses and preprocessed data files for this paper are available at https://tinyurl.com/detecting-social-info] Data from videos were binned according to the age of the child (in weeks). First, we saw that the proportion of faces in view showed a moderate decrease across this age range (see Figure \ref{fig:faceHandDetailed}), in keeping with prior findings [@Fausey2016]; in contrast, we did not observe an increase in the proportion of hands in view. These effects were quantified with two separate linear mixed-models (see Tables 1 & 2).^[Face/hand detections were binned across each week of filming. Participant's age was converted into months and centered for these analyses. Random slopes for the effect of age by child led to a singular fit and were removed from both analyses; see full model specification in accompanying codebase.]

However, the most striking result from these analyses is a much overall greater proportion of hands in view than have previously been reported [@Fausey2016]. We found this to be true across all ages, in all three children, and regardless of whether we analyzed human annotations (on the 24K random subset, see dotted lines in Appendix Figure \ref{fig:goldSetSanity}) or OpenPose annotations on the entire dataset (see Figure \ref{fig:faceHandDetailed}A). This is notable especially given that OpenPose showed relatively low recall for hands, indicating that this may be an underestimate of the proportion of hands in view. In fact, analysis the human annotations underscores revealed a much higher proportion of hands relative to faces than the automated annotations.

@@

<!-- FIXME - references not rendering for -->
As we found that children's hands tended to occur in the lower visual field (see Figure \ref{fig:handLocation}), we thus re-analyzed the entire dataset while restricting our analysis to the center field of view, decreasing the proportion of hand detections from `r round(sum(all_vid_data$num_hands)/sum(all_vid_data$num_frames),2)*100`% to `r round(sum(all_vid_data$num_hands_center,na.rm=TRUE)/sum(all_vid_data$num_frames),2)*100`%, but only decreased face detections from `r round(sum(all_vid_data$num_faces)/sum(all_vid_data$num_frames),2)*100`% to `r round(sum(all_vid_data$num_faces_center,na.rm=TRUE)/sum(all_vid_data$num_frames),3)*100`%. This cropping likely removed both the majority of detections of children's own hands but also some detections of adult hands (see Figure \ref{fig:handLocation}), especially as OpenPose was biased to miss children's hands when they were in view. Nonetheless, within this modified field of view, we still observed more hand detections than face detections (see dashed lines in Figure \ref{fig:faceHandDetailed}). We also still found a higher proportion of hands in view relative to faces when excluding any frames containing child hand's from the human annotated gold sample (see Appendix Figure\ref{fig:goldSetSanity}).

Finally, we analyzed how these two sources of social information co-occurred, finding that faces/hands were jointly present in `r sum(all_vid_data$faces_and_hands,na.rm=TRUE)/sum(all_vid_data$num_frames)*100` percent of frames (see face hand-occurrences across age in Figure \ref{fig:faceHandDetailed}C). To do so, we  calculated the number of frames in which infants saw faces and hands together relative to overall proportions of faces/hands that were detected for each child and age range. We found that all three infants than were more likely see hands independently -- without the presence of a face -- that they were likely to see faces independently \ref{fig:faceHandContingnecy}. That is, generally speaking when a face was present, a hand also tended to be present.

As we found that children's hands tended to occur in the lower visual field (see Figure \ref{fig:handLocation}), we thus re-analyzed the entire dataset while restricting our analysis to the center field of view, decreasing the proportion of hand detections from `r round(sum(all_vid_data$num_hands)/sum(all_vid_data$num_frames),2)*100`% to `r round(sum(all_vid_data$num_hands_center,na.rm=TRUE)/sum(all_vid_data$num_frames),2)*100`%, but only decreased face detections from `r round(sum(all_vid_data$num_faces)/sum(all_vid_data$num_frames),2)*100`% to `r round(sum(all_vid_data$num_faces_center,na.rm=TRUE)/sum(all_vid_data$num_frames),3)*100`%. 
This cropping likely removed both the majority of detections of children's own hands but also some detections of adult hands (see Figure \ref{fig:handLocation}), especially as OpenPose was biased to miss children's hands when they were in view. 
Nonetheless, within this modified field of view, we still observed more hand detections than face detections (see dashed lines in Figure \ref{fig:faceHandDetailed}B). 
We also still found a higher proportion of hands in view relative to faces when excluding any frames containing child hand's from the human annotated gold sample (see Appendix Figure\ref{fig:goldSetSanity}).

@@

@@

Finally, we also examined whether there were changes in where faces tended to appear in the camera's (and hence, by proxy, the child's) field of view. 
As expected, faces tended to be located towards the upper field of view, while views of hands were more centrally distributed (see Appendix, Figure \ref{fig:faceHandDensity} for average density distributions). 
However, we also found evidence that older children tended to see more faces in more variable positions than younger children. 
Specifically, we examined how variable the horizontal and vertical coordinates were of the faces in the infant view. 
To do so, we calculated the coefficient of variation of the horizontal (x) and vertical (y) positions of centers of the faces detected by OpenPose (see Figure \ref{fig:faceDetailed}B), and examined changes across age. 
Faces tended to be more variable in the vertical than their horizontal position (see Figure \ref{fig:faceDetailed}B). 
We also found that as children got older, they tended to see faces that varied more in their horizontal -- but not their vertical position -- suggesting that older children might be more likely to see more smaller faces in their periphery (see Figure \ref{fig:faceDetailed}B).
